---
title: "Tidy Tuesday - U.S. Customs and Border Protection Encounters Dataset"
subtitle: "Tuesday Number 3"
author: "ALEC CHEN"
format: 
  html:
    embed-resources: true
editor: 
  markdown: 
    wrap: 72
---
#### Preface

This dataset tracks how many people CBP encountered trying to enter the U.S. through different borders and methods, capturing both Title 8 (standard immigration law) and Title 42 (COVID-era health policy) cases from 2020 onwards to help understand migration patterns and enforcement changes.

## Research Question

> Which top fragrance notes are most frequently associated with the leading fragrance brands, and how are these top brands and top notes interconnected?

## Data Wrangling

```{r}
library(tidyverse)
library(lubridate)
library(plotly)
library(htmlwidgets)

cbp_resp <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-26/cbp_resp.csv')
cbp_state <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-26/cbp_state.csv')

month_map <- c(
  JAN = 1, FEB = 2, MAR = 3, APR = 4, MAY = 5, JUN = 6,
  JUL = 7, AUG = 8, SEP = 9, OCT = 10, NOV = 11, DEC = 12
)

# Convert fiscal_year and month_abbv into a date.
# Treat fiscal_year as if it matches the calendar year.
cbp_resp_dates <- cbp_resp %>%
  mutate(
    month_num = month_map[month_abbv],
    year_month_date = make_date(year = fiscal_year, month = month_num, day = 1)
  ) %>%
  # Keep only Title 8 and Title 42 data
  filter(title_of_authority %in% c("Title 8", "Title 42"))
```

Conglomerating data

```{r}
# Aggregate data by month and title_of_authority
monthly_data <- cbp_resp_dates %>%
  group_by(title_of_authority, year_month_date) %>%
  summarise(encounters = sum(encounter_count), .groups = "drop") %>%
  arrange(title_of_authority, year_month_date) %>%
  group_by(title_of_authority) %>%
  mutate(
    lag_encounters = lag(encounters),
    pct_change = (encounters - lag_encounters) / lag_encounters * 100
  ) %>%
  ungroup()

# Create tooltip text
monthly_data <- monthly_data %>%
  mutate(hover_text = paste0(
    "Date: ", format(year_month_date, "%Y-%m"), "<br>",
    "Encounters: ", encounters, "<br>",
    "MoM Change: ", ifelse(is.na(pct_change), "N/A", paste0(round(pct_change, 1), "%"))
  ))

# Draw Milestones lines
policy_milestones <- tibble(
  event = c("Title 42 Implemented", "Title 42 End"),
  date = as.Date(c("2020-03-21", "2023-05-11"))
)
```


## Data Visualization 

**Static Plot of encounters**

```{r}
# First calculate max y value
max_y <- max(monthly_data$encounters)

# Generate base plot
base_encounters_plot <- ggplot(monthly_data, aes(x = year_month_date, 
                             y = encounters, 
                             group = title_of_authority,
                             text = hover_text)) +
  geom_line(aes(color = title_of_authority), size = 1) +
  scale_x_date(date_breaks = "6 months", date_labels = "%Y-%m") +
  # Reset inherited aesthetics with inherit.aes = FALSE
  geom_text(
    data = policy_milestones,
    mapping = aes(x = date,
                 y = max_y),
    inherit.aes = FALSE,  # This is the key addition
    label = policy_milestones$event,
    color = "red",
    hjust = -0.1,
    vjust = 0.5,
    size = 3.5,
    angle = 0
  ) +
  geom_vline(data = policy_milestones,
             aes(xintercept = date),
             color = "red",
             linetype = "dashed") + 
  labs(
    title = "Title 8 & Title 42 Encounters Over Time",
    x = "Date",
    y = "Number of Encounters",
    color = "Authority"
  ) +
  scale_color_manual(
    values = c("Title 42" = "#FF9999", "Title 8" = "#66C2E5")
  ) +
  scale_y_continuous(labels = scales::comma_format()) +
  theme_classic() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )

ggsave("CBP_encounters.png", plot = base_encounters_plot, bg = 'white', width = 12, height = 6, dpi = 300)

base_encounters_plot
```

**Interactive plot of encounters**

```{r}
# Create the plot directly with plotly for both lines
interactive_encounters_plot <- plot_ly() %>%
  add_trace(data = monthly_data[monthly_data$title_of_authority == "Title 8", ],
            x = ~year_month_date, 
            y = ~encounters,
            name = "Title 8",
            type = 'scatter',
            mode = 'lines',
            line = list(color = "#66C2E5", width = 2)) %>%
  add_trace(data = monthly_data[monthly_data$title_of_authority == "Title 42", ],
            x = ~year_month_date,
            y = ~encounters,
            name = "Title 42",
            type = 'scatter',
            mode = 'lines',
            line = list(color = "#FF9999", width = 2)) %>%
  layout(
    title = "Title 8 & Title 42 Encounters Over Time",
    xaxis = list(title = "Date"),
    yaxis = list(title = "Number of Encounters"),
    shapes = list(
      # Title 42 Implementation line
      list(
        type = "line",
        x0 = "2020-03-21",
        x1 = "2020-03-21",
        y0 = 0,
        y1 = max(monthly_data$encounters),
        line = list(color = "red", width = 2, dash = "dash")
      ),
      # Title 42 End line
      list(
        type = "line",
        x0 = "2023-05-11",
        x1 = "2023-05-11",
        y0 = 0,
        y1 = max(monthly_data$encounters),
        line = list(color = "red", width = 2, dash = "dash")
      )
    ),
    annotations = list(
      # Title 42 Implementation text
      list(
        x = "2020-03-21",
        y = max(monthly_data$encounters),
        text = "Title 42 Implemented",
        showarrow = FALSE,
        xanchor = "left",
        yanchor = "bottom",
        font = list(color = "red")
      ),
      # Title 42 End text
      list(
        x = "2023-05-11",
        y = max(monthly_data$encounters),
        text = "Title 42 End",
        showarrow = FALSE,
        xanchor = "left",
        yanchor = "bottom",
        font = list(color = "red")
      )
    ),
    hovermode = "x unified"
  )

htmlwidgets::saveWidget(interactive_encounters_plot , "CBP_interactive_encounters_plot.html")
```


#### Exmaple plot

Red line in plotly

```{r}
library(plotly)

# Example dataframe
df <- data.frame(
  date = seq(as.Date("2020-01-01"), as.Date("2024-01-01"), by = "month"),
  value = runif(49, 100, 300)
)

# Figuring out how to put red lines onto the plot
plot_ly(df, x = ~date, y = ~value, type = 'scatter', mode = 'lines') %>%
  layout(
    shapes = list(
      list(
        type = "line",
        x0 = "2022-01-01",
        x1 = "2022-01-01",
        y0 = 100,
        y1 = 300,
        line = list(color = "red", width = 2, dash = "dash")
      )
    )
  )
```


