---
title: "Tidy Tuesday - U.S. Customs and Border Protection Encounters Dataset"
subtitle: "Tuesday Number 2"
author: "ALEC CHEN"
format: 
  html:
    embed-resources: true
editor: 
  markdown: 
    wrap: 72
---
#### Preface

This dataset tracks how many people CBP encountered trying to enter the U.S. through different borders and methods, capturing both Title 8 (standard immigration law) and Title 42 (COVID-era health policy) cases from 2020 onwards to help understand migration patterns and enforcement changes.

## Research Question

> Which top fragrance notes are most frequently associated with the leading fragrance brands, and how are these top brands and top notes interconnected?

## Data Wrangling

```{r}
library(tidyverse)
library(lubridate)
library(plotly)

cbp_resp <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-26/cbp_resp.csv')
cbp_state <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-26/cbp_state.csv')

month_map <- c(
  JAN = 1, FEB = 2, MAR = 3, APR = 4, MAY = 5, JUN = 6,
  JUL = 7, AUG = 8, SEP = 9, OCT = 10, NOV = 11, DEC = 12
)

# Convert fiscal_year and month_abbv into a date.
# Treat fiscal_year as if it matches the calendar year.
cbp_resp_dates <- cbp_resp %>%
  mutate(
    month_num = month_map[month_abbv],
    year_month_date = make_date(year = fiscal_year, month = month_num, day = 1)
  ) %>%
  # Keep only Title 8 and Title 42 data
  filter(title_of_authority %in% c("Title 8", "Title 42"))
```

Conglomerating data

```{r}
# Aggregate data by month and title_of_authority
monthly_data <- cbp_resp_dates %>%
  group_by(title_of_authority, year_month_date) %>%
  summarise(encounters = sum(encounter_count), .groups = "drop") %>%
  arrange(title_of_authority, year_month_date) %>%
  group_by(title_of_authority) %>%
  mutate(
    lag_encounters = lag(encounters),
    pct_change = (encounters - lag_encounters) / lag_encounters * 100
  ) %>%
  ungroup()

monthly_data <- monthly_data %>%
  mutate(hover_text = paste0(
    "Date: ", format(year_month_date, "%Y-%m"), "<br>",
    "Encounters: ", encounters, "<br>",
    "MoM Change: ", ifelse(is.na(pct_change), "N/A", paste0(round(pct_change, 1), "%"))
  ))

# Draw Milestones lines
policy_milestones <- tibble(
  event = c("Title 42 Implemented", "Title 42 End"),
  date = as.Date(c("2020-03-21", "2023-05-11"))
)
```

```{r}
ggplot(monthly_data, aes(x = year_month_date, y = encounters, color = title_of_authority)) +
  geom_line(size = 1) +
  geom_vline(data = policy_milestones, 
             aes(xintercept = date),
             color = "red", 
             linetype = "dashed") +
  # Improved text annotations with better positioning and formatting
  geom_text(
    data = policy_milestones,
    aes(x = date, 
        y = max(monthly_data$encounters),  # Position at top of plot
        label = event),
    color = "red", 
    hjust = -0.1,  # Adjust horizontal position
    vjust = 0.5,   # Center vertically
    size = 3.5,    # Slightly larger text
    angle = 0      # Horizontal text instead of vertical
  ) +
  labs(
    title = "Title 8 & Title 42 Encounters Over Time",
    x = "Date",
    y = "Number of Encounters",
    color = "Authority"
  ) +
  # Format y-axis to use comma notation instead of scientific
  scale_y_continuous(labels = scales::comma_format()) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
