---
title: "Tidy Tuesday - College Admissions Dataset"
subtitle: "Tuesday Number 1"
author: "ALEC CHEN"
format: 
  html:
    embed-resources: true
editor: 
  markdown: 
    wrap: 72
---

#### Preface

For my first TidyTuesday in two years! (can't believe it's been that long), I have to bring myself back to the COMP 112 days and the [Claus Wilke's fundamentals](https://clauswilke.com/dataviz/)

I must try to implement some of these ideas that he posits

- *Providing Context*: Balance the data and the context
- *Proportional Ink*: The sizes of shaded areas in a visualization need to be proportional to the data values they represent.
- *Redundant encoding*: If there is a clear visual ordering in your data, make sure to match it in the legend. 
- *Avoiding 3D or line drawings*: Just don't.

Also, trying some fun techniques for handling overlapping data like: Partial transparency and jittering, 2D histograms, and contour lines.

## Research Question

I mean, generally we know that students that come from higher economic statuses have more resources, are admitted at much higher rate, and generally do better in life.

But is there an instance where money is detrimental to career and life success? Thus my inital research question was: 

> Where and why do higher income students fail in college admissions?

After close inspection of the data, I don't think I can visualize this research question. The data looks at college attendance which is very clearly tied to higher income [study](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10089351/). Since Macalester College is in this dataset, I will change my research question:  

> How does Macalester College compare to other universities in terms of college attendance across economic brackets?

## Exploring the Data

Load the data

```{r}
library(tidyverse)
library(hrbrthemes)
library(plotly)
college_admissions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-09-10/college_admissions.csv')
dim(college_admissions)
head(college_admissions)
```
#### Preliminary Visualizations

Single Variable visualizations: It looks like much of the data cleaning work has already been done for me. The hard part is understanding the data dictionary

```{r, collapse=TRUE}
# This shouldn't happen
ggplot(college_admissions) + 
  geom_bar(aes(x = par_income_lab))

# Odd that they put the same number of people in each bracket, probably for visualization's sake
table(college_admissions$par_income_lab)

# Of every income bracket, around 0-0.1 actually attend a specific college. Ex: around 1% of students in 0-20 income bracket attend Cornell (weighting all test scores equal)
ggplot(college_admissions) + 
  geom_histogram(aes(x = attend))

ggplot(college_admissions) + 
  geom_histogram(aes(x = rel_apply_unwgt))

# Of all income brackets, around 1-4% of those people actually apply to a singular school (ie UCLA)
ggplot(college_admissions) + 
  geom_histogram(aes(x = rel_apply))

# Lots of selective colleges in this dataset
ggplot(college_admissions) +
  geom_bar(aes(x = tier)) + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) + # Using a function in labeling is really cool
  theme(axis.text.x = element_text(hjust = 0.5))

# Flagship, Ivy Plus, and NESCAC are over-represented in this dataset
ggplot(college_admissions) +
  geom_bar(aes(x = test_band_tier)) + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme(axis.text.x = element_text(hjust = 0.5))
```

Two Variable Visualizations:

```{r, collapse=TRUE}
# This looks like it should be a boxplot
ggplot(college_admissions) +  
  geom_point(aes(x = par_income_bin, y = rel_attend))

# Use relative attendance because it's normalized between colleges
ggplot(college_admissions) + 
  geom_boxplot(aes(x = par_income_lab, y = rel_attend))

# Macalester ain't doing too bad!
college_admissions %>%
  filter(name == "Macalester College") %>%
  ggplot() +
    geom_col(aes(x = par_income_lab, y = attend)) # geom_col is like barplot but for data that's better

other_elite_schools <- college_admissions %>%
  filter(tier == "Other elite schools (public and private)") %>% 
  count(name)
```

Given Grinnell's list of near peer colleges, I list some near peer colleges for Macalester that I will use in this visualization

1. Amherst College
2. Bowdoin College
3. Carleton College
4. Davidson College
5. Kenyon College
6. Oberlin College
7. Pomona College
8. Reed College
9. Swarthmore College
10. Vassar College
11. Williams College

I chose these colleges because they all are rigorous academically and small liberal arts.

#### Full Visualization

```{r}
#full_plot <- 
  
  
college_admissions %>%
  filter(name %in% c("Amherst College", "Bowdoin College", "Carleton College", "Davidson College", "Kenyon College", "Macalester College", "Oberlin College", "Pomona College", "Reed College", "Swarthmore College", "Vassar College", "Williams College")) %>% 
  mutate(name = as.factor(name)) %>%  # Ensure name is a factor
  mutate(name = fct_reorder(name, rel_attend, .fun = mean, .desc = TRUE)) %>%
  ggplot() + 
    geom_tile(aes(x = par_income_lab, y = name, fill = rel_attend)) + 
    labs(title = "How economically diverse are liberal arts campuses?", subtitle = "Relative Attendances across all Income Brackets", x = "Income Brackets", y = "Liberal Arts Colleges", fill = "Relative Attendance", caption = "Alec Chen | Data from Opportunity Insights") + 
    scale_fill_viridis_c(option = "magma",labels = function(x) paste0(x, "%")) + 
    #scale_fill_distiller(palette = "Spectral") +
    guides(fill = guide_colorbar(title.position = "top", barwidth = 15, barheight = 1)) +  # Moves title to top
    scale_x_discrete(labels = c(
    "0-20" = "0-20%", "20-40" = "20-40%", "40-60" = "40-60%", 
    "60-70" = "60-70%", "70-80" = "70-80%", "80-90" = "80-90%",
    "90-95" = "90-95%", "95-96" = "95-96%", "96-97" = "96-97%", 
    "97-98" = "97-98%", "98-99" = "98-99%", "99-99.9" = "99-99.9%", 
    "Top 0.1" = "Top 0.1%", "Top 1" = "Top 1%"
    )) + 
    theme_classic() +
    theme(
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 10, hjust = 0.5),
      legend.position='top', 
      legend.justification='left',
      legend.direction='horizontal',
      legend.margin = margin(b = -10, l = 5, unit = "pt"),
      # axis.text.x = element_text(size = 11, angle = 45, hjust = 1),  # Adjusts x-axis labels for readability
      axis.text.x = element_text(size = 11),
      axis.text.y = element_text(size = 11, hjust = 1),
      axis.title.y = element_blank(),
      axis.title.x = element_text(size = 12),
      axis.ticks = element_blank(),
      axis.line = element_blank())
```

Save Plot

```{r}
ggsave("chen_economic_diversity_colleges_tidy_tuesday_1.png", plot = full_plot, width = 12, height = 6, dpi = 300)
```


