---
title: "Tidy Tuesday - Perfume Dataset"
subtitle: "Tuesday Number 2"
author: "ALEC CHEN"
format: 
  html:
    embed-resources: true
editor: 
  markdown: 
    wrap: 72
---
#### Preface

This dataset examines how perfumes are rated, identifies their associated scent families, and highlights the brands and perfumers responsible for creating top fragrances.

## Research Question

> Which top fragrance notes are most frequently associated with the leading fragrance brands, and how are these top brands and top notes interconnected?

## Data Wrangling

Load Packages

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(igraph)
library(ggraph)
library(ggplot2)
library(tidyverse)
```

Load the data

```{r}
parfumo_data_clean <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-12-10/parfumo_data_clean.csv')
```

Identify top notes and brands. (note mean scent in this context)

```{r}
# Converts top notes into a character vector
top_notes <- brand_note_counts %>%
  filter(Note != "note_NA") %>%
  group_by(Note) %>%
  summarise(total_freq = sum(Freq)) %>% #select the top 20 notes by total frequency
  top_n(20, total_freq) %>%
  pull(Note)

# Converts top brands into a character vector
top_brands <- brand_note_counts %>%
  filter(Note != "brand_NA") %>%
  group_by(Brand) %>%
  summarise(total_freq = sum(Freq)) %>%
  top_n(20, total_freq) %>%
  pull(Brand)
```

Get entries with top 20 brands and top 20 notes

```{r}
filtered_edges <- brand_note_counts %>%
  filter(Note %in% top_notes & Brand %in% top_brands)
```

Set up the graph - source node: brand - target node: Note - edge weight: Frequency

```{r}
# Convert to graph with prefixes
filtered_edges <- filtered_edges %>%
  mutate(from = paste0("brand_", Brand),
         to = paste0("note_", Note),
         weight = Freq) %>%
  select(from, to, weight)
```

Create distinct brand and note nodes.

```{r}
nodes_brand <- filtered_edges %>%
  distinct(from) %>%
  rename(name = from) %>%
  mutate(type = TRUE)  # TRUE = Brand

nodes_note <- filtered_edges %>%
  distinct(to) %>%
  rename(name = to) %>%
  mutate(type = FALSE) # FALSE = Note

nodes <- bind_rows(nodes_brand, nodes_note)
```

Create a graph from filtered_edges

```{r}
g <- graph_from_data_frame(d = filtered_edges, vertices = nodes, directed = FALSE)
```

## Visualization

Visualize the graph

```{r}
p <- ggraph(g, layout = "fr") + # force-directed graph 
  geom_edge_link(aes(width = weight, color = weight), alpha = 0.4) + # Draws edges
  geom_node_point(aes(color = type), size = 4) + # Draws nodes
  geom_node_text(aes(label = name), repel = TRUE, size = 3) + # Adds labels
  scale_color_manual(values = c("TRUE" = "darkred", "FALSE" = "darkgreen"),
                     labels = c("Brand", "Note")) + # Set colors
  scale_edge_width(range = c(0.5, 2)) +
  scale_edge_color_gradient(low = "skyblue", high = "darkblue") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(title = "Brand vs. Top Notes Network (Top 20)",
       caption = "Data From Parfumo",
       edge_color = "Frequency",
       edge_width = "Frequency")

ggsave("Brand_Top_Note_network.png", plot = p, bg = white, width = 12, height = 6, dpi = 300)

print(p)
```

> Analysis: The dense cluster of edges shows that the top brands share many top notes in common. This could reflect which industry trend about which top notes are popular or trendy. The top notes that are closer to the center like Bergamo are likely "universal favorites" that are popular regardless of the time.
